Robust Clustering for Ad Hoc Cognitive Radio Network

### 2017-04-01 journal-draft/bare_jrnl_comsoc.tex
message compelxity

We give the definition of cluster in CRN as follows. 
A cluster $C$ is a set of secondary nodes which possess the same set of common channels.
In particular, a cluster consists of a cluster head $h(C)$ and a number of cluster members, and the cluster head is able to communicate with any cluster member directly.
A cluster can be formed only by the cluster head.
%$\text{Nb}(i)$ denotes node $i$'s neighborhood which consists of all its one hop neighbors.
%For any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
The size of $C$ is denoted by $|C|$.
When the cluster head of a cluster is $i$, we denote that cluster by $C(i)$.
$K(C)$ denotes the set of CCs in cluster $C$, $ K(C) = \Cap_{i\in C} K_i$.
%Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model are listed in Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & set of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h(C)$ & the cluster head of a cluster C\\
%$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
% $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
 $C_i$  & the $i$-th cluster in $\mathcal{S}$ \\
 $|C_i|$ & size of the cluster $C_i$\\
 $|K(C_i)|$ & the number of CCs of cluster $C_i$\\
\bottomrule
\end{tabular}
\end{table}

@@

In fact, the number of steps which are actually involved in this process is much smaller than $N^3$, as both $n$ and $m$ are considerably smaller than $N$.
The percentage of debatable nodes in $\mathcal{N}$ is illustrated in Figure~\ref{percentage_overlapping_node}, which is between 10\% to 60\% of the total number of CR nodes in the network.
The number of clusters heads, as mentioned in Section~\ref{phaseI}, is dependent on the network density and the CR node's transmission range.
The simulation in \cite{robust_clustering_arxiv} shows that the cluster heads are only 3.4\% to 20\% of the total CR nodes.

@@

\subsubsection{Control Signaling Overhead}
%Different from the clustering schemes proposed in ~\cite{LIU_TMC11_2, clustering_globecom11}, 
%There are two phases for any variants of ROSS.
%Clusters are formed in the first phase, in the second phase, cluster membership is decided so that each node only resides in one cluster.
%Control message exchanges between CR nodes are involved in both phases.

In this section we compare the overhead of signaling involved in different clustering schemes.
%In order to highlight the amount of control signalling only for clustering, 
We don't consider the the control messages which are involved in neighborhood discovery, which is the premise and deemed to be the same for all clustering schemes.
According to~\cite{complexity_aggregation_2011}, the message complexity is defined as the number of messages used by all nodes.
To have the same metric to compare, we count \textit{the number of transmissions of control messages}, without distinguishing broadcast or uni-cast control messages.
This metric is synonymous with \textit{the number of updates} discussed in Section~\ref{ross}.

As to ROSS, the control messages are generated in both phases.
In the first phase, when a CR node decides itself to be the cluster head, it broadcasts a message containing its ID, cluster members and the set of CCs in its cluster.
In the second phase, a debatable node broadcasts its affiliation to inform its claiming clusters, then the cluster heads of the claiming clusters broadcast message about the new cluster members if they are changed due to the debatable node's decision.
The upper bound of the total number of the control messages involved in cluster formation is analyzed in Theorem~\ref{clustering:theorem} and Section~\ref{clustering:phaseII:game}.

The comparison scheme SOC consists of three rounds of execution. 
In the first two rounds, every CR node maintains its own cluster and seeks either to integrate neighboring clusters or to join one neighboring cluster.
The final clusters are obtained in the third round. 
In each round, every CR node is involved in comparisons and cluster mergers.
%Comparing with the second phase of ROSS, only debatable nodes to communicate with cluster heads to clarify their membership.
%The signalling overhead for centralized scheme comes from two processes, the the process of collecting information to the centralized controller, and the process that the controller spreads the clustering result to all the CR nodes.

The centralized scheme is conducted at the centralized control device, but it involves two phases of control message transmission.
The first phase is information aggregation, in which every CR node's channel availability and neighborhood is transmitted to the centralized controller.
iIn the second phase, the control broadcasts the clustering solution, which is disseminated to every CR node.
%
We adopt the algorithm proposed in~\cite{Efficient_broadcasting_gathering_adhoc} to broadcast and gather information as the algorithm is simple and self-stabilizing.
This scheme needs building a backbone structure to support the communication. 
We apply ROSS to generate cluster heads which serve as the backbone, and the debatable nodes are used as the gateway nodes between the backbone nodes.
As the backbone is built for one time and supports the transmission of control messages later on, we don't take account the messages involved in building the backbone.
As to the process of information gathering, we assume that every cluster member sends the spectrum availability and its ID to its cluster head, which further forwards the message to the controller, then the number of transmissions is $N$.
As to the process of dissemination, in an extreme situation where all the gateway and the backbone nodes broadcast, the number of transmissions is $h + m$, where $h$ is the number of cluster heads and $m$ is number of debatable nodes.
%, $d$ is the average number of demanding clusters for each debatable node.

The number of control messages which are involved in ROSS variants and the centralized scheme is related with the number of debatable nodes.
Figure~\ref{percentage_overlapping_node} shows the percentage of debatable nodes with different network densities, from which we can obtain the value of $m$.
\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.6\linewidth]{percentage_overlapping_node.pdf}
  \caption{The percentage of debatable nodes after phase I of ROSS.}\label{percentage_overlapping_node}
\end{figure}
%
%As we adopt a simplified communication model, node's transmission is not influenced by collision or interference, 
%Our simulation doesn't consider the behaviour in the physical layer, 
%
%
%Assume we use OSPF~\cite{BCJ10} to aggregate and disseminate information, then the best and worst complexity for is $\mathcal{O}(E)$, where $E$ is the number of edges in the graph which corresponds to the network.
%The minimum number of edges is $n-1$ when the nodes form a line and each node has at more two neighbours, and the maximum number is $N*(N-1)/2$ when the nodes form a complete graph.
%Thus the best message complexity of the centralized scheme is $\mathcal{O}{(N)}$ and the worst is $\mathcal{O}{(N^2)}$.
%
%1. update membership to form X1, 
%2. broadcast new X1, form new X2
%3. broadcast X3
%The complexity parameters are the number of nodes $n$ in network, number of clusters $h$.
Table \ref{tab_overhead} shows the message complexity, quantitative amount of the control messages, and the size of control messages.
Figure~\ref{control_msg} shows the analytical result of the amount of transmissions involved in different schemes.
%the upper bound of the number of transmissions of ROSS, and the analytic number of transmissions of the centralized scheme.

@@

\begin{center}
\begin{table*}[!htb]
\caption{Signalling overhead}\label{tab_overhead}
{\renewcommand{\arraystretch}{1.15} %<- modify value to suit your needs
{\small
\hfill{}
\begin{threeparttable}
\begin{tabular}{|C{2 cm}|C{2.8 cm}|C{3.15 cm}|C{7.7 cm}|}
\hline
 Scheme 				&Message Complexity 	&   Quantitative number of messages 		& Content of message (size of message) 									\\ \hline
 ROSS-DGA, ROSS-$\delta$-DGA 	&$\mathcal{O}(N^3)$ (worst case)		&   $h+2m^2d$ (upper bound)  				&   \multirow{2}{*}{\parbox{7.7cm}{Cluster head $i$ broadcasts channel availability on all cluster members ($|C(i)| |\mathcal{K}|$ bytes); Cluster member $i$ broadcasts the new individual connectivity $d_i$ after being included in one or more clusters (1 byte)}}								\\ \cline{1-3}
 ROSS-DFA, ROSS-$\delta$-DFA 	&$\mathcal{O}(N)$ (worst case)		&   $h + 2m$	 (upper bound) 					& 	      												\\ \hline
 SOC 					&$\mathcal{O}(N)$		&   $3N$									& Every CR node $i$ broadcasts channel availability on all cluster members ($|C(i)| |\mathcal{K}|$ bytes)
 \\ \hline
 Centralized			&$\mathcal{O}(N)$			&	$h + m + N$ (upper bound)~\cite{Efficient_broadcasting_gathering_adhoc}		& clustering result (2$N$ bytes) \tnotex{tnote:robots-r1} 					\\ \hline
\end{tabular}
    \begin{tablenotes}
      \item\label{tnote:robots-r1}Assuming the data structure of the clustering result is in the form of \{ Node ID $i$, cluster head ID h($C$) where $i\in C$, for every $i\in \mathcal{N}$ \}.
    \end{tablenotes}
    \end{threeparttable}
}
}
\hfill{}
\end{table*}
\end{center}

@@

\subsection{Comparison among the Distributed Schemes}
\label{largeScaleCRN}
In this section we investigate the performances of distributed clustering schemes in CRN with different network scales and densities.
The transmission range of CR is $A/5$, PU's transmission range is $2A/5$.
The initial number of PU is 30.
%The number of CR is 100, 200 and 300, and the average number of neighbours of each CR is 9.5, 20, and 31.
The desired sizes adopted are listed in the Table~\ref{Simulation_para}, which is about 60\% of the average number of neighbours.
%When run ROSS, the parameter $t$ which is used to control cluster size in phase I is 1.3.

@@

\subsubsection{Robustness of the clusters against newly added PUs}
We add extra 20 batches of PUs sequentially in the CRN, where each batch includes 10 PUs. 
Figure~\ref{singleton_clusters_100} and \ref{singleton_clusters_200} show that when $N=100$ and 200, more unclustered CR nodes appear in the CRN where SOC is applied.
When the network becomes denser, as shown in Figure~\ref{singleton_clusters_300}, ROSS-DGA/DFA generate slightly more unclustered CR nodes than SOC when new PUs are not many, but SOC's performance deteriorates quickly when the number of PUs becomes larger.
We only show the average values of the variants of ROSS as their confidence intervals overlap.
%
When applying ROSS with size control mechanism, significantly less unclustered CR nodes are generated.
Besides, the greedy mechanism moderately strengthens the robustness of the clusters.

>>>>

We give the definition of cluster in CRN as follows. 
A cluster $C$ is a set of secondary nodes which possess the same set of common channels.
In particular, a cluster consists of a cluster head $h(C)$ and a number of cluster members, and the cluster head is able to communicate with any cluster member directly.
A cluster can be formed only by the cluster head.
%$\text{Nb}(i)$ denotes node $i$'s neighborhood which consists of all its one hop neighbors.
%For any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
The size of $C$ is denoted by $|C|$.
When the cluster head of a cluster is $i$, we denote that cluster by $C(i)$.
$K(C)$ denotes the set of CCs in cluster $C$, $ K(C) = \Cap_{i\in C} K_i$.
%Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model are listed in Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & set of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h(C)$ & the cluster head of a cluster C\\
%$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
% $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
 $C_i$  & the $i$-th cluster in $\mathcal{S}$ \\
 $|C_i|$ & size of the cluster $C_i$\\
 $|K(C_i)|$ & the number of CCs of cluster $C_i$\\
 $n$ & the number of debatable nodes\\
 $m$ & the number of claiming cluster heads\\
\bottomrule
\end{tabular}
\end{table}

@@

In fact, the upper bound of the number of steps which are actually involved in this process is much smaller than $N^3$.%, as both $n$ and $m$ are considerably smaller than $N$.
The percentage of debatable nodes in the network is illustrated in Figure~\ref{percentage_overlapping_node}, which is between 10\% to 60\% of the total number of CR nodes in the network.
In the other hand, the number of clusters heads is dependent on the network density and the CR node's transmission range as mentioned in Section~\ref{phaseI}.
The simulation in \cite{robust_clustering_arxiv} shows that the cluster heads are only 3.4\% to 20\% of the total CR nodes.

@@

\subsubsection{Control Signaling Overhead}

In this section we compare the overhead of signaling involved in different clustering schemes.
We count the number of \textit{transmissions of control messages} as message complexity~\cite{complexity_aggregation_2011}, and without distinguishing broadcast or uni-cast control messages.
In Section~\ref{ross}, this metric is synonymous with the \textit{the number of updates}.
%We don't consider the the control messages which are involved in neighborhood discovery, which is the premise and deemed to be the same for all clustering schemes.
%According to~\cite{complexity_aggregation_2011}, the message complexity is defined as the number of messages used by all nodes.
%To have the same metric to compare,

As to ROSS, in the first phase the maximal number of broadcast is $N$ according to~\ref{clustering:theorem}.
The upper bounds of the transmissions are $n^2m$ and $n$ for ROSS-DGA and ROSS-DFA respectively.
Scheme SOC consists of three rounds and in each round every node needs to broadcast in order to do comparisons and cluster mergers.
The centralized scheme is conducted at the centralized control device, and it involves information aggregation and clustering decision dissemination.
We adopt the backbone structure proposed in~\cite{Efficient_broadcasting_gathering_adhoc} to analyze the centralized scheme's message complexity.
We apply ROSS to generate cluster heads which serve as the backbone.
In the process of information aggregation, all the nodes transmit information to the cluster heads which forward the messages to the controller, then in the process of dissemination, all the cluster heads and the debatable nodes broadcast the clustering result, thus the upper bound of the number of broadcast is $N+m+n$.

%
%, when a CR node decides itself to be the cluster head, it broadcasts a message containing its ID, cluster members and the set of CCs in its cluster.
%In the second phase, a debatable node broadcasts its affiliation to inform its claiming clusters, then the cluster heads of the claiming clusters broadcast message about the new cluster members if they are changed due to the debatable node's decision.
%The upper bound of the total number of the control messages involved in cluster formation is analyzed in Theorem~\ref{clustering:theorem} and Section~\ref{clustering:phaseII:game}.
%
%The comparison scheme SOC consists of three rounds of execution. 
%In the first two rounds, every CR node maintains its own cluster and seeks either to integrate neighboring clusters or to join one neighboring cluster.
%The final clusters are obtained in the third round. 
%In each round, every CR node is involved in comparisons and cluster mergers.
%%Comparing with the second phase of ROSS, only debatable nodes to communicate with cluster heads to clarify their membership.
%%The signalling overhead for centralized scheme comes from two processes, the the process of collecting information to the centralized controller, and the process that the controller spreads the clustering result to all the CR nodes.
%
%The centralized scheme is conducted at the centralized control device, but it involves two phases of control message transmission.
%The first phase is information aggregation, in which every CR node's channel availability and neighborhood is transmitted to the centralized controller.
%iIn the second phase, the control broadcasts the clustering solution, which is disseminated to every CR node.
%%
%We adopt the algorithm proposed in~\cite{Efficient_broadcasting_gathering_adhoc} to broadcast and gather information as the algorithm is simple and self-stabilizing.
%This scheme needs building a backbone structure to support the communication. 
%We apply ROSS to generate cluster heads which serve as the backbone, and the debatable nodes are used as the gateway nodes between the backbone nodes.
%As the backbone is built for one time and supports the transmission of control messages later on, we don't take account the messages involved in building the backbone.
%As to the process of information gathering, we assume that every cluster member sends the spectrum availability and its ID to its cluster head, which further forwards the message to the controller, then the number of transmissions is $N$.
%As to the process of dissemination, in an extreme situation where all the gateway and the backbone nodes broadcast, the number of transmissions is $m + n$, where $m$ is the number of cluster heads and $n$ is number of debatable nodes.

The number of control messages which are involved in ROSS variants and the centralized scheme is related with the number of debatable nodes.
Figure~\ref{percentage_overlapping_node} shows the percentage of debatable nodes with different network densities.
\begin{figure}[ht!]
  \centering
  \includegraphics[width=0.6\linewidth]{percentage_overlapping_node.pdf}
  \caption{The percentage of debatable nodes after phase I of ROSS.}\label{percentage_overlapping_node}
\end{figure}
%
%As we adopt a simplified communication model, node's transmission is not influenced by collision or interference, 
%Our simulation doesn't consider the behaviour in the physical layer, 
%
%
%Assume we use OSPF~\cite{BCJ10} to aggregate and disseminate information, then the best and worst complexity for is $\mathcal{O}(E)$, where $E$ is the number of edges in the graph which corresponds to the network.
%The minimum number of edges is $n-1$ when the nodes form a line and each node has at more two neighbours, and the maximum number is $N*(N-1)/2$ when the nodes form a complete graph.
%Thus the best message complexity of the centralized scheme is $\mathcal{O}{(N)}$ and the worst is $\mathcal{O}{(N^2)}$.
%
%1. update membership to form X1, 
%2. broadcast new X1, form new X2
%3. broadcast X3
%The complexity parameters are the number of nodes $n$ in network, number of clusters $h$.
Table \ref{tab_overhead} shows the message complexity, quantitative amount of the control messages, and the size of control messages.
Figure~\ref{control_msg} shows the analytical result of the amount of transmissions involved in different schemes.
%the upper bound of the number of transmissions of ROSS, and the analytic number of transmissions of the centralized scheme.

@@

\begin{center}
\begin{table*}[!htb]
\caption{Signalling overhead}\label{tab_overhead}
{\renewcommand{\arraystretch}{1.15} %<- modify value to suit your needs
{\small
\hfill{}
\begin{threeparttable}
\begin{tabular}{|C{2 cm}|C{2.8 cm}|C{3.15 cm}|C{7.7 cm}|}
\hline
 Scheme 				&Message Complexity 	&   Quantitative number of messages 		& Content and size of the message 									\\ \hline
 ROSS-DGA, ROSS-$\delta$-DGA 	&$\mathcal{O}(N^3)$ (worst case)		&   $N+n^2m$ (upper bound)  				&   \multirow{2}{*}{\parbox{7.7cm}{PhaseI: ID, $d_i$,$g_i$, which are 3 bytes;\\ PhaseII: Cluster head $i$ broadcasts channel availability on all cluster members, where are $|C(i)| |\mathcal{K}|$ bytes}}								\\ \cline{1-3}
 ROSS-DFA, ROSS-$\delta$-DFA 	&$\mathcal{O}(N)$ (worst case)		&   $N + n$	 (upper bound) 					& 	      												\\ \hline
 SOC 					&$\mathcal{O}(N)$		&   $3N$									& Every CR node $i$ broadcasts channel availability on all cluster members ($|C(i)| |\mathcal{K}|$ bytes)
 \\ \hline
 Centralized			&$\mathcal{O}(N)$			&	$N + n + m$ (upper bound) \cite{Efficient_broadcasting_gathering_adhoc}		& clustering result, which is 2$N$ bytes \tnotex{tnote:robots-r1} 					\\ \hline
\end{tabular}
    \begin{tablenotes}
      \item\label{tnote:robots-r1}Assuming the data structure of the clustering result is in the form of $\{i, C\}, i\in C, i\in \mathcal{N}$.
    \end{tablenotes}
    \end{threeparttable}
}
}
\hfill{}
\end{table*}
\end{center}

@@

\subsection{Comparison among the Distributed Schemes}
\label{largeScaleCRN}
In this section we investigate the performances of the proposed distributed clustering schemes in CRN with different network scales and densities.
The transmission range of CR is $A/5$, PU's transmission range is $2A/5$.
The initial number of PU is 30.
%The number of CR is 100, 200 and 300, and the average number of neighbours of each CR is 9.5, 20, and 31.
The desired sizes adopted are listed in the Table~\ref{Simulation_para}, which is about 60\% of the average number of neighbours.
%When run ROSS, the parameter $t$ which is used to control cluster size in phase I is 1.3.

@@

\subsubsection{Robustness of the clusters against increased PUs}
We add extra 20 batches of PUs sequentially in the CRN, where each batch includes 10 PUs. 
Figure~\ref{singleton_clusters_100} and \ref{singleton_clusters_200} show that when $N=100$ and 200, more unclustered CR nodes appear in the CRN where SOC is applied.
When the network becomes denser, as shown in Figure~\ref{singleton_clusters_300}, ROSS-DGA/DFA generate slightly more unclustered CR nodes than SOC when new PUs are not many, but SOC's performance deteriorates quickly when the number of PUs becomes larger.
We only show the average values of the variants of ROSS as their confidence intervals overlap.
%
When applying ROSS with size control mechanism, significantly less unclustered CR nodes are generated.
Besides, the greedy mechanism moderately strengthens the robustness of the clusters.
